---
title: "lasso_Sig_vs_Raw"
author: "hong"
date: "3/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list=ls())
setwd("~/Desktop/TD_Proj_Data")


### Sig. Data ###
sig.adduct <- read.csv("2_Integration/Data/sig.adduct.csv")
sig.adduct$X <- NULL

sig.metab <- read.csv("2_Integration/Data/sig.metab.csv")
sig.metab$X <- NULL

sig.miRNA <- read.csv("2_Integration/Data/sig.miRNA.csv")
sig.miRNA$X <- NULL

### Raw Data ###
adductomics <- read.csv("2_Integration/Data/matched_adducts.csv")  
adductomics$X <- NULL

metabolites <- read.csv("2_Integration/Data/matched_metabolites.csv")
metabolites$X <- NULL
metabolites <- as.data.frame(scale(metabolites))

miRNA <- read.csv("2_Integration/Data/matched_miRNA.csv")
miRNA$X <- NULL
miRNA = as.data.frame(scale(miRNA))
```


############## Lasso Model ##################



1. Adducts[Y] vs. Metabolites[X] 

Context:
You can have a pollutant that happens to go through "metabolic activation", so this pollutant is now a metabolite. Now, can this metabolite bind DNA? If so, it is a DNA-adduct. 

Hypothesis:
a) "Are there sets of metabolites significantly correlated with adducts that were significantly correlated with TRAP?"
b) So then, are any of these metabolites also adducts? - ie. are these metabolites capable of binding DNA, thus, becoming adducts?
c) This is looking at a separate pathway. Could the appearance of sign. adducts be a result of metabolites?
  ci) Could the appearance of sign. adducts be a result of sign. metabolites (related to exp)?

```{r}
library(lme4)

#glmnet expects matrix of predictors
metabolites <- as.matrix(metabolites)
adductomics <- as.matrix(adductomics)
miRNA <- as.matrix(miRNA)

###
set.seed(1)
lassomodels.1se = lassomodels = NULL

t0 = Sys.time()
for (k in 1:ncol(sig.adduct)) {
  print(k)
  model.lasso = cv.glmnet(x = metabolites, y = sig.adduct[,k])
  lassomodels.1se = cbind(lassomodels.1se, coef(model.lasso, s = "lambda.1se")[-1])
  lassomodels = cbind(lassomodels, coef(model.lasso, s = "lambda.min")[-1])
}
t1 = Sys.time()
print((t1-t0))

colnames(lassomodels) = colnames(lassomodels.1se) = colnames(sig.adduct)
rownames(lassomodels) = rownames(lassomodels.1se) = colnames(metabolites)

saveRDS(lassomodels.1se, "2_Integration/Results/Lasso/adduct.metab[X]lasso_1se.rds")
saveRDS(lassomodels, "2_Integration/Results/Lasso/adduct.metab[X]lasso_min.rds")
```

```{r} {Number of metabolites[X] related to each sig.adduct[Y]}

add.metab_1se = readRDS("2_Integration/Results/Lasso/adduct.metab[X]lasso_1se.rds")
add.metab.min = readRDS("2_Integration/Results/Lasso/adduct.metab[X]lasso_min.rds")

#lambda w/ 1SE
apply(add.metab_1se, 2, FUN = function(x) {
  sum(x != 0)
})

#lambda.min
apply(add.metab.min, 2, FUN = function(x) {
  sum(x != 0)
})
```

1b) sig.Metab[Y] vs. Adducts[X] (Very Secondary)
- Is there a set of adducts that is linearly significant to sig.Metab? 
- Implication: Are there DNA defects/modifications that leads to different levels of metabolites we see (related to exposure)? 
- Need to consider the layers between adducts-->metabolites
```{r}
set.seed(1)
lassomodels.1se = lassomodels = NULL

t0 = Sys.time()
for (k in 1:ncol(sig.metab)) {
  print(k)
  model.lasso = cv.glmnet(x = adductomics, y = sig.metab[,k])
  lassomodels.1se = cbind(lassomodels.1se, coef(model.lasso, s = "lambda.1se")[-1])
  lassomodels = cbind(lassomodels, coef(model.lasso, s = "lambda.min")[-1])
}
t1 = Sys.time()
print((t1-t0))

colnames(lassomodels) = colnames(lassomodels.1se) = colnames(sig.metab)
rownames(lassomodels) = rownames(lassomodels.1se) = colnames(adductomics)

saveRDS(lassomodels.1se, "2_Integration/Results/Lasso/metab.adduct[X]lasso_1se.rds")
saveRDS(lassomodels, "2_Integration/Results/Lasso/metab.adduct[X]lasso_min.rds")
```


c) sig.adduct[Y] vs. sig.metab[X]
- of the sig.metab are there linear combinations metabs correlated to individual sig. adducts?

- probably doesn't make sense to run sig.metab[Y] vs. sig.adduct[X], because can adducts lead to different levels of metabolites?
```{r}
sig.metab <- as.matrix(sig.metab)

set.seed(1)
lassomodels.1se = lassomodels = NULL

t0 = Sys.time()
for (k in 1:ncol(sig.adduct)) {
  print(k)
  model.lasso = cv.glmnet(x = sig.metab, y = sig.adduct[,k])
  lassomodels.1se = cbind(lassomodels.1se, coef(model.lasso, s = "lambda.1se")[-1])
  lassomodels = cbind(lassomodels, coef(model.lasso, s = "lambda.min")[-1])
}
t1 = Sys.time()
print((t1-t0))

colnames(lassomodels) = colnames(lassomodels.1se) = colnames(sig.adduct)
rownames(lassomodels) = rownames(lassomodels.1se) = colnames(sig.metab)

saveRDS(lassomodels.1se, "2_Integration/Results/Lasso/sigAdduct.sigMetab[X]lasso_1se.rds")
saveRDS(lassomodels, "2_Integration/Results/Lasso/sigAdduct.sigMetab[X]lasso_min.rds")
```

2) sig.metabolites[Y] vs. miRNA[X]
- are there silencing artifacts that led to the metabolites we see related to exp?

*probably not much significance running sig.miRNA[Y] vs. metabolites[X] - can metabolites affect miRNA levels? Unlikely, unless the metabolites are DNA/RNA-adducts

```{r}
set.seed(1)
lassomodels.1se = lassomodels = NULL

t0 = Sys.time()
for (k in 1:ncol(sig.metab)) {
  print(k)
  model.lasso = cv.glmnet(x = miRNA, y = sig.metab[,k])
  lassomodels.1se = cbind(lassomodels.1se, coef(model.lasso, s = "lambda.1se")[-1])
  lassomodels = cbind(lassomodels, coef(model.lasso, s = "lambda.min")[-1])
}
t1 = Sys.time()
print((t1-t0))

colnames(lassomodels) = colnames(lassomodels.1se) = colnames(sig.metab)
rownames(lassomodels) = rownames(lassomodels.1se) = colnames(miRNA)

saveRDS(lassomodels.1se, "2_Integration/Results/Lasso/metab.miRNA[X]lasso_1se.rds")
saveRDS(lassomodels, "2_Integration/Results/Lasso/metab.miRNA[X]lasso_min.rds")
```

3. sig.miRNA[Y] vs. Adducts[X]
- can modifications to DNA lead to differential expression of miRNA? - Perhaps.

**(sig.adduct[Y] vs. miRNA[X])
- can RNA silencing lead to modification in DNA? - Unlikely.

```{r}
set.seed(1)
lassomodels.1se = lassomodels = NULL

t0 = Sys.time()
for (k in 1:ncol(sig.miRNA)) {
  print(k)
  model.lasso = cv.glmnet(x = adductomics, y = sig.miRNA[,k])
  lassomodels.1se = cbind(lassomodels.1se, coef(model.lasso, s = "lambda.1se")[-1])
  lassomodels = cbind(lassomodels, coef(model.lasso, s = "lambda.min")[-1])
}
t1 = Sys.time()
print((t1-t0))

colnames(lassomodels) = colnames(lassomodels.1se) = colnames(sig.miRNA)
rownames(lassomodels) = rownames(lassomodels.1se) = colnames(adductomics)

saveRDS(lassomodels.1se, "2_Integration/Results/Lasso/miRNA.adduct[X]lasso_1se.rds")
saveRDS(lassomodels, "2_Integration/Results/Lasso/miRNA.adduct[X]lasso_min.rds")
```
