---
title: "Integrate_Univariate_Sig.vs.Raw"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

#UNIVARIATE ANALYSIS OF SIGNIFICANT RESULTS


```{r}
rm(list=ls())
setwd("~/Desktop/TD_Proj_Data/")

#Raw Data
load("~/Desktop/TD_Proj_Data/Oxford_Street_Full_peaks_imputed_metabolomics.RData")
metabolites <- Full_peaks_imputed

adductomics <- read.csv("Adductomics_Oxford_Street_all_imputed.csv")

miRNA_data <- data.frame(t(read.table("OXF2_Tx_miRNA_procd.txt")))
miRNA_target <- read.csv("OXF2_Tx_miRNA_TargetFile.txt", sep="")

#Read in Significant Omics
multiomic_sig <- read.csv("1_GLS_Results/multiomic_significant_vars.csv")
multiomic_sig$d_143.1.1 <- NULL #removing duplicate OMIC
###multiomic_sig <- multiomic_sig[-c(53/54),] delete either 53 or 54

#Subsetting Significant results into distinct Omics
sig.adduct <- multiomic_sig[,c(2,433:434)] #2613 duplicate
sig.adduct <- sig.adduct[-c(54),]


sig.metab <- multiomic_sig[,c(284:432)] #2613 is duplicated
sig.metab <- sig.metab[-c(54),]

sig.miRNA <- multiomic_sig[,c(3:8)] #2613 is duplicated
sig.miRNA <- sig.miRNA[-c(54),]

####Issue w/ mRNA
#sig.mRNA <- multiomic_sig[,c(2, 9:283)] #2613 has different values
######


#Making Raw have the same number of rows as significant set
### For miRNA

#Adding OMIC.ID to raw miRNA
#Altering miRNA rownames
rownames <- rownames(miRNA_data)
rownames <- gsub("[.]", "-", rownames)
rownames(miRNA_data) <- rownames

#Building index for miRNA
miRNA_data$SampleCode <- rownames(miRNA_data) 
miRNA <- inner_join(miRNA_target, miRNA_data, by = "SampleCode")

miRNA$ID <- as.numeric(substring(miRNA$SampleCode,3,5))
miRNA$Time_Point <- as.factor(ifelse(grepl("4pm", miRNA$SampleCode), "T2 (4PM)", 
                                     ifelse(grepl("8am", miRNA$SampleCode),"T1 (8AM)","T3 (24hrs)")))
#hyde park = 1, Oxford =2
# treatment 1-3 = HP, treatment 4-6 = Oxford

numbers <- ifelse(miRNA$Treatment == 1, 11, 
                  ifelse(miRNA$Treatment == 2, 12, 
                         ifelse(miRNA$Treatment == 3, 13, 
                                ifelse(miRNA$Treatment == 4, 21, 
                                       ifelse(miRNA$Treatment == 5, 22,23)))
                  )
)

miRNA$XP_ID <- paste(miRNA$ID,numbers, sep = "")
miRNA <- miRNA[,c(373:375,1:372)] 
miRNA <- within(miRNA, rm("hsa.miR.451a"))
rownames(miRNA) <- miRNA$XP_ID

#Include only miRNA Omics
miRNA <- miRNA[,10:ncol(miRNA)]

#Creating index for matching OMIC.ID
select.idx <- multiomic_sig$XP_ID
select.idx <- select.idx[-c(54)]

#Matching to raw miRNA
miRNA <- miRNA[intersect(select.idx, rownames(miRNA)),]

dir.create("2_Integration/Data", showWarnings = FALSE)
write.csv(miRNA, file = "2_Integration/Data/matched_miRNA.csv")


############### For Metabolites ######################

rownames(metabolites) <- metabolites$XP_ID
metabolites <- metabolites[intersect(select.idx, rownames(metabolites)), ]
metabolites <- metabolites[,7:ncol(metabolites)]
metabolites$XP_ID <- NULL

write.csv(metabolites, file = "2_Integration/Data/matched_metabolites.csv")

################ For Adduct ########################

#Problem: 29 is duplicated, need to find which duplicated is right by matching exp. values
#Exp. values don't match Sig. values

#Finding which batch in raw corresponds to sig.adduct
add_29 <- adductomics[74:85,]
log_add_29 <- log(add_29[,18:ncol(add_29)])
log_add_29$XP_ID <- add_29$XP_ID
log_add_29$id <- add_29$id
log_add_29 <- log_add_29[,c(33,34,1:32)]
#Discard later batch belonging to person_29
adductomics <- adductomics[-c(79,75,77,85,81,83),]

rownames(adductomics) <- adductomics$XP_ID
adductomics <- adductomics[intersect(select.idx, rownames(adductomics)), ]
adductomics <- adductomics[,18:ncol(adductomics)]
adductomics <- log(adductomics)

write.csv(adductomics, file = "2_Integration/Data/matched_adducts.csv")

#################### Univariate Model #########################
#X - entire omic set
#Y - significant omics

##### Loading the data ######

#Significant Set
sig.miRNA = as.data.frame(scale(sig.miRNA))
sig.metab = as.data.frame(scale(sig.metab))

sig.adduct = as.data.frame(scale(sig.adduct))
sig.adduct$XP_ID <- NULL

#Whole Set
miRNA = read.csv("2_Integration/Data/matched_miRNA.csv")
miRNA$X <- NULL
miRNA = as.data.frame(scale(miRNA))

metabolites <- read.csv("2_Integration/Data/matched_metabolites.csv")
metabolites$X <- NULL
metabolites <- sapply(metabolites, as.numeric)
metabolites <- scale(metabolites)
metabolites <- as.data.frame(metabolites)

adductomics <- read.csv("2_Integration/Data/matched_adducts.csv")
adductomics$X <- NULL
adductomics <- as.data.frame(scale(adductomics))


#### MODEL #####


########################### sig.miRNA ############################

### vs. Metabolies[X] ###
Beta = Pvalues = matrix(NA, nrow = ncol(sig.miRNA), ncol=ncol(metabolites))

t0 = Sys.time()
for (i in 1:ncol(sig.miRNA)) {
  for (j in 1:ncol(metabolites)) {
    model1 = lm(sig.miRNA[, i] ~ metabolites[, j])
    Beta[i, j] = coefficients(model1)["metabolites[, j]"]
    Pvalues[i, j] = summary(model1)$coefficients["metabolites[, j]", "Pr(>|t|)"]
  }
}
rownames(Pvalues) = rownames(Beta) = colnames(sig.miRNA)
colnames(Pvalues) = colnames(Beta) = colnames(metabolites)


saveRDS(Pvalues, "2_Integration/Results/miRNA.metab[X]_univ_pval.rds")
saveRDS(Beta, "2_Integration/Results/miRNA.metab[X]_univ_betas.rds")


### vs. Adducts [X] ###
Beta = Pvalues = matrix(NA, nrow = ncol(sig.miRNA), ncol=ncol(adductomics))

t0 = Sys.time()
for (i in 1:ncol(sig.miRNA)) {
  for (j in 1:ncol(adductomics)) {
    model1 = lm(sig.miRNA[, i] ~ adductomics[, j])
    Beta[i, j] = coefficients(model1)["adductomics[, j]"]
    Pvalues[i, j] = summary(model1)$coefficients["adductomics[, j]", "Pr(>|t|)"]
  }
}
rownames(Pvalues) = rownames(Beta) = colnames(sig.miRNA)
colnames(Pvalues) = colnames(Beta) = colnames(adductomics)

saveRDS(Pvalues, "2_Integration/Results/miRNA.adduct[X]_univ_pval.rds")
saveRDS(Beta, "2_Integration/Results/miRNA.adduct[X]_univ_betas.rds")



#################### sig.Metabolites ####################

### vs. miRNA[X] ###

Beta = Pvalues = matrix(NA, nrow = ncol(sig.metab), ncol=ncol(miRNA))

t0 = Sys.time()
for (i in 1:ncol(sig.metab)) {
  for (j in 1:ncol(miRNA)) {
    model1 = lm(sig.metab[, i] ~ miRNA[, j])
    Beta[i, j] = coefficients(model1)["miRNA[, j]"]
    Pvalues[i, j] = summary(model1)$coefficients["miRNA[, j]", "Pr(>|t|)"]
  }
}
t1 = Sys.time()
print(t1-t0)

rownames(Pvalues) = rownames(Beta) = colnames(sig.metab)
colnames(Pvalues) = colnames(Beta) = colnames(miRNA)

saveRDS(Pvalues, "2_Integration/Results/metab.miRNA[X]_univ_pval.rds")
saveRDS(Beta, "2_Integration/Results/metab.miRNA[X]_univ_betas.rds")


### vs. Adductomics[X] ###

Beta = Pvalues = matrix(NA, nrow = ncol(sig.metab), ncol=ncol(adductomics))

t0 = Sys.time()
for (i in 1:ncol(sig.metab)) {
  for (j in 1:ncol(adductomics)) {
    model1 = lm(sig.metab[, i] ~ adductomics[, j])
    Beta[i, j] = coefficients(model1)["adductomics[, j]"]
    Pvalues[i, j] = summary(model1)$coefficients["adductomics[, j]", "Pr(>|t|)"]
  }
}
t1 = Sys.time()
print(t1-t0)

rownames(Pvalues) = rownames(Beta) = colnames(sig.metab)
colnames(Pvalues) = colnames(Beta) = colnames(adductomics)

saveRDS(Pvalues, "2_Integration/Results/metab.adduct[X]_univ_pval.rds")
saveRDS(Beta, "2_Integration/Results/metab.adduct[X]_univ_betas.rds")



######################### sig.Adduct ########################## 

### vs. miRNA[X] ###

Beta = Pvalues = matrix(NA, nrow=ncol(sig.adduct), ncol=ncol(miRNA))

t0 = Sys.time()
for (i in 1:ncol(sig.adduct)) {
  print(i)
  for (j in 1:ncol(miRNA)) {
    model1 = lm(sig.adduct[, i] ~ miRNA[, j])
    Beta[i, j] = coefficients(model1)["miRNA[, j]"]
    Pvalues[i, j] = summary(model1)$coefficients["miRNA[, j]", "Pr(>|t|)"]
  }
}
t1 = Sys.time()
print(t1-t0)

rownames(Pvalues) = rownames(Beta) = colnames(sig.adduct)
colnames(Pvalues) = colnames(Beta) = colnames(miRNA)

saveRDS(Pvalues, "2_Integration/Results/adduct.miRNA[X]_univ_pval.rds")
saveRDS(Beta, "2_Integration/Results/adduct.miRNA[X]_univ_betas.rds")



### vs. metabolites[X] ###

Beta = Pvalues = matrix(NA, nrow=ncol(sig.adduct), ncol=ncol(metabolites))

t0 = Sys.time()
for (i in 1:ncol(sig.adduct)) {
  print(i)
  for (j in 1:ncol(metabolites)) {
    model1 = lm(sig.adduct[, i] ~ metabolites[, j])
    Beta[i, j] = coefficients(model1)["metabolites[, j]"]
    Pvalues[i, j] = summary(model1)$coefficients["metabolites[, j]", "Pr(>|t|)"]
  }
}
t1 = Sys.time()
print(t1-t0)

rownames(Pvalues) = rownames(Beta) = colnames(sig.adduct)
colnames(Pvalues) = colnames(Beta) = colnames(metabolites)

saveRDS(Pvalues, "2_Integration/Results/adduct.metabolites[X]_univ_pval.rds")
saveRDS(Beta, "2_Integration/Results/adduct.metabolites[X]_univ_betas.rds")
```



Finding Sig. Pairs according to 'BH'

```{r}
add.metab_pval <- readRDS("2_Integration/Results/adduct.metabolites[X]_univ_pval.rds")
add.miRNA_pval <- readRDS("2_Integration/Results/adduct.miRNA[X]_univ_pval.rds")

miRNA.metab_pval <- readRDS("2_Integration/Results/miRNA.metab[X]_univ_pval.rds")
miRNA.add_pval <- readRDS("2_Integration/Results/miRNA.adduct[X]_univ_pval.rds")

metab.miRNA_pval <- readRDS("2_Integration/Results/metab.miRNA[X]_univ_pval.rds")
metab.add_pval <- readRDS("2_Integration/Results/metab.adduct[X]_univ_pval.rds")

pval_list <- list(add.metab_pval, add.miRNA_pval, miRNA.add_pval, miRNA.metab_pval, metab.miRNA_pval, metab.add_pval)

#Showing # of sig. pairs in each OMICS comparison
for (i in pval_list) {
  print(sum(p.adjust(as.vector(i), method = "bonf") < 0.05))
}
# Significant Pairs after BH: miRNA[Y].metab[X] = 129, metab.miRNA[X] = 216, metab.add[X]= 3